Legacy Data Confounds Modern Genomics Studies

### 2019-03-25 PlottingScripts/ProcessPlotData.Rmd
update

#save to file
saveRDS(HanKgpAF,
            paste0(DataPath,
                   'HanKgpAF.rds'))
```

##Full Model
```{r}
#read Genotypes
han1=fread(paste0(DataPath,
                  'SignificantSNPs_Han90_0.01_genotype.txt'))
HanNames=fread(paste0(DataPath,
                      'SignificantSNPs_Han90_COLNAMES.txt'),
               header=F)
names(han1) = as.character(unlist(HanNames))
#melt to long format
han = melt(han1, id=c('#CHROM','POS','ID'))
#fix chromosome names
han$CHROM = as.numeric(as.character(substring(han$'#CHROM',4)))
han$'#CHROM' <- NULL

#to deal with INDELs we only count one instance of duplicate positions (i.e. pick max GT)
han <- han %>% 
  group_by(CHROM,POS,variable) %>% 
  summarize(han.gt = max(value))
#to get allele frequencies we group by POS and summarize GT
hanAF <- han %>% 
  group_by(CHROM,POS) %>% 
  summarize(hanAF = sum(han.gt))

#read Genotypes
kgp1 = fread(paste0(DataPath,
                    'SigVar_Pop_0.01_genotype.txt'))
ColNames = read.table(paste0(DataPath,
                             'ColNames.txt'),
                      header=F)
names(kgp1) = as.character(unlist(ColNames))
#melt to long format
kgp = melt(kgp1, id=c('CHROM','POS','ID'))
#keep only the 83 resequenced individuals
kgp = kgp[which(kgp$variable %in% han$variable),]

#to deal with INDELs we only count one instance of duplicate positions (i.e. pick max GT)
kgp <- kgp %>% 
  group_by(CHROM,POS,variable) %>% 
  summarize(Kgp.gt = max(value))
#to get allele frequencies we group by POS and summarize GT
kgpAF <- kgp %>% 
  group_by(CHROM,POS) %>% 
  summarize(kgpAF = sum(Kgp.gt))
#keep only positions that are present in the 83 individuals
kgpAF <- kgpAF[which(kgpAF$kgpAF > 0),]

#merge 1kGP and han resequenced data.
combo <- merge(hanAF, kgpAF, by=c('CHROM','POS'), all.y=T)
#missing sites replaced with 0
combo[is.na(combo)] <- 0

##read Indels
Indels <- fread(paste0(DataPath,
                    'Reg_indels.POS'), 
                skip=1 ,col.names=c('CHROM','POS'))
Indels$Indel = 'yes'

INDEL = merge(Indels, combo, by=c('CHROM','POS'))
SNP = combo[which((combo$POS %in% INDEL$POS)==F),]

#save to file
saveRDS(INDEL,
            paste0(DataPath,
                   'INDEL_HanKgp.rds'))
#save to file
saveRDS(SNP,
            paste0(DataPath,
                   'SNP_HanKgp.rds'))

SNP
INDEL
```
##Extract Significant Positions from 1kGP Data
```{bash}
DataPath='/Users/luke/Documents/QualityPaper/ProcessedData'
RawDataPath='/Users/luke/genomes/genomes/hg19/phase3'
VCF='.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz'

@@

```
##Compare Over and Under Qual 30
```{r}
data_bim <- fread(paste0(DataPath,
                         'Significant_Positions.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)

>>>>

#save to file
saveRDS(HanKgpAF,
            paste0(DataPath,
                   'HanKgpAF.rds'))
```
##Extract Significant Positions from 1kGP Data
```{bash}
DataPath='/Users/luke/Documents/QualityPaper/ProcessedData'
RawDataPath='/Users/luke/genomes/genomes/hg19/phase3'
VCF='.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz'

@@

```

```{bash}
DataPath='/Users/luke/Documents/QualityPaper/ProcessedData'
RawDataPath='/Users/luke/genomes/genomes/Han90/90_Han_Chinese_SNP_gatkvqsr_20170424_chr'
VCF='_validator.vcf.gz'

#Han vcf has `chr` before chromosome number
sed -e 's/^/chr/' \
$DataPath/Significant_Positions.txt \
> $DataPath/Significant_Positions_chr.txt

#GetHeader
zgrep '^#' $RawDataPath\22$VCF > $DataPath/Significant_Positions_Han.vcf

#AddGenotypes
for f in `seq 1 22`;
do echo $f;
bcftools view \
-R $DataPath/Significant_Positions_chr.txt \
$RawDataPath${f}$VCF \
| grep -v "^#" \
>> $DataPath/Significant_Positions_Han.vcf ; \
done

#bcftools doesn't take rsID
#this deals with indels and other mulitallelic variants
vcftools \
--vcf $DataPath/Significant_Positions_Han.vcf \
--snps $DataPath/Significant_rsID.txt \
--recode --recode-INFO-all \
--out $DataPath/Significant_Positions_Han

/Users/luke/bin/plink_mac/plink \
--vcf $DataPath/Significant_Positions_Han.recode.vcf \
--make-bed \
--out $DataPath/Significant_Positions_Han
```

##Full Model
```{r}
sharedID <- fread(paste0(DataPath,
                          '83samples.txt'), 
                  header = FALSE,
                  col.names = 'sample')

kgp_bim <- fread(paste0(DataPath,
                         'Significant_Positions.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)

kgp_fam <- fread(paste0(DataPath,
                         'Significant_Positions.fam'), 
                  col.names = c('familyID','sample',
                                'father','mother',
                                'sex','pheno')) %>%
  select(sample)

kgp_genotype <- read.bed(paste0(DataPath,'Significant_Positions'))
rownames(kgp_genotype) <- kgp_bim$rsID
colnames(kgp_genotype) <- kgp_fam$sample
#Only keep the 83 shared samples
keep <- colnames(kgp_genotype)[(colnames(kgp_genotype) 
                                 %in% sharedID$sample)]
kgp_genotype <- kgp_genotype[,keep]
kgp_genotype <- kgp_genotype[!duplicated(rownames(kgp_genotype)),]
kgp_genotype[1:5,1:5]

kgp <- melt(kgp_genotype, 
           varnames = c('rsID','sample'), 
           value.name = 'Genotype')
kgp <- 
		kgp %>% 
		group_by(rsID) %>% 
		summarize(kgp.AF = sum(Genotype))

kgp <- kgp[which(kgp$kgp.AF > 0),]
#read Han Genotypes
han_bim <- fread(paste0(DataPath,
                         'Significant_Positions_Han.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)

han_fam <- fread(paste0(DataPath,
                         'Significant_Positions_Han.fam'), 
                  col.names = c('familyID','sample',
                                'father','mother',
                                'sex','pheno')) %>%
  select(sample)

han_genotype <- read.bed(paste0(DataPath,'Significant_Positions_Han'))
rownames(han_genotype) <- han_bim$rsID
colnames(han_genotype) <- han_fam$sample
#Only keep the 83 shared samples
keep <- colnames(han_genotype)[(colnames(han_genotype) 
                                 %in% sharedID$sample)]
han_genotype <- han_genotype[,keep]
han_genotype <- han_genotype[!duplicated(rownames(han_genotype)),]
han_genotype[1:5,1:5]

han = melt(han_genotype, 
           varnames = c('rsID','sample'), 
           value.name = 'Genotype')

han <- 
		han %>% 
		group_by(rsID) %>% 
		summarize(han.AF = sum(Genotype))

#merge 1kGP and han resequenced data.
combo <- merge(han, kgp, by='rsID', all.y=T)
#missing sites replaced with 0
combo[is.na(combo)] <- 0

#save to file
saveRDS(combo,
            paste0(DataPath,
                   'HanKgp_AF.rds'))

combo
```

##Compare Over and Under Qual 30
```{r}
data_bim <- fread(paste0(DataPath,
                         'Significant_Positions.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)
